---
title: "covid 19"
author: "Nicole Orizino"
date: "Spring 2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex

$H_0$: Covid 19 deaths in Korea, Jam-May 2020 are unrelated 
$H_A$: There is a relation between deaths in Korea Jan-May 2020 and sex of the patient


Table of patients:
```{r}
kRdata$Sex = as.factor(kRdata$Sex)
kRdata$result = as.factor(kRdata$result)


# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

This chart shows that more males died from covid than female, with a 25 for female and 44 for male.

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

`
```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=Sex,y=count,fill=result))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=Sex,y=prcnt*100,fill=result)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  

The gg plot count graph shows that females who caught Covid survived more than males, with a count of around 2200 for female and 1400 for male, while more males died from covid than female, with a count of about 25 female and 50 male. 

The second gg plot, the percent plot shows that about 55% of females survived while 45% of males survived. And 45% females died and 65% males died from covid.  

Both graphs show that the female sex in Korea survived more that the male sex of those who caught Covid 19.


## Numerical results


```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The Numerical results also shows that more females lived while more males died. 56f vs 44m alive and 36f vs 64m dead.


## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


The P-value is .001 and in all the results (observed,expected by Null and chi-square statistics) shows that females survived covid more than men. Resulting in Reject the Null Hypothesis. 
The odds ratio also shows males are 2.3 more times to die due to covid.


## Conclusions

The question was is dying independent of sex? The conclusion to this question is, we reject the null hypothesis. The P-value is .001 meaning it is a small number. And all the graphs here show that more females survived covid 19 than male. 
